### 2019-04-04 analysis/mssel_analysis.Rmd
stuck on XPEHH results

---
title: "Mssel_analysis"
author: "Patrick Monnahan"
date: "12/27/2018"
output: html_document
---

## Load packages and import data
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(assertthat)
library(ggplot2)
library(dplyr)
library(tidyr)
library(PopGenome)
library(stringr)
library(data.table)
library(magrittr)
library(wrapr)

ms2 = read.table("~/Documents/Research/PloidySim/data/mssel_out_2-24-19_fin.txt", head = T)

#Group windows into larger sets
ms2 %<>% mutate(ints = as.numeric(as.character(cut(bp.start, breaks = 50, labels = seq(1,50)))))

```

#Background

## Dominance scalar
This plot shows the dominance value of each genotype (given by within individual allele frequency) as a funciton of the dominance scalar
```{r}
getDomCoefs = function(dominance, ploidy = 100){
  if (dominance == 0){
    coeffs = rep(0, ploidy - 1)
  }
  else if (dominance == 1){
    coeffs = rep(1, ploidy - 1)
  }
  else {
    dom = (1 - dominance) - 0.5
    dom = abs((dom * 10 + sign(dom)) ^ sign(dom))
    coeffs = (seq(1, ploidy - 1) / ploidy) ^ dom
  }
  return(coeffs)
}

doms = c(seq(0,1, 0.1))
Doms = as.data.frame(sapply(doms, getDomCoefs))
Doms = rbind(c(0,0,0,0,0,0,0), Doms, c(1,1,1,1,1,1,1))
colnames(Doms) = doms
Doms = cbind(Doms, data.frame("i/k"=seq(0,100) / 100))
mDoms = melt(Doms, id.var = "i.k")
ggplot(mDoms, aes(x = i.k, y = value, color = variable)) + geom_line() + scale_color_discrete(name = "Dominance\nScalar") + theme_bw() + xlab("i / k") + ylab("Dominance Coefficient")
```

## Allele frequency trajectories based on deterministic recursion equations
```{r}
#Function creates trajectories
getTraj = function(s, dom, ploidy, start_freq, N = -9, end_freq = 0.99, maxGens = 9999, maxTries = 10){
  dom_coeffs = getDomCoefs(dom, ploidy)
  attempts = 0
  p = start_freq
  k = ploidy
  gen = 0
  while (attempts <= maxTries & p < end_freq & gen <= maxGens){
    df = data.frame("s" = s, "dom" = dom, "gen" = 0, "freq" = start_freq, "dp1" = 0, "w.bar" = 0, "ploidy" = ploidy)
    p = start_freq
    dp1 = 0
    gen = 0
    het_fits = 1 + dom_coeffs * s
    fits = c(1, het_fits, 1 + s) / (1 + s)
    while (p < end_freq & p > 0 & gen <= maxGens){
      q = 1 - p
      i = seq(0, k)
      Num = sum(choose(k, i) * ((k - i) / k) * (p ^ (k - i)) * (q ^ i) * rev(fits))
      Den = sum(choose(k, i) * (p ^ (k - i)) * (q ^ i) * rev(fits))
      p_prime = Num / Den
      if (N != -9){
        p_prime = sum(rbinom(N, k, p_prime)) / (k * N) 
      }
      df = rbind(df, c(s, dom, gen, p, dp1, Den, k))
      dp1 = p_prime - p
      p = p_prime
      gen = gen + 1
    }
    df = rbind(df, c(s, dom, gen, p, dp1, Den, k))
    attempts = attempts + 1
  }
  if (attempts > maxTries & nrow(df) <= maxGens){
    print(paste("Beneficial allele was lost due to drift for", maxTries, "consecutive attempts"))
  }
  if (nrow(df) > maxGens){
    print(paste("Beneficial allele did not fix before the maximum number of generations (", maxGens, ")."))
  }
  return(df[-1,])
}

# Param set to create trajectories for
params = data.frame("s" = rep(0.1,9), "dom" = c(rep(0.5,3), rep(0,3), rep(1,3)), "ploidy" = rep(c(2,4,8),3), "start" = rep(0.05, 9))

# Create trajectories
dat1 = foreach(i = 1:nrow(params), .combine=rbind) %do% {
  getTraj(params[i,]$s, params[i,]$dom, params[i,]$ploidy, params[i,]$start)
}

#plot
ggplot(dat1, aes(x=gen, y=freq, linetype=as.factor(dom), color=as.factor(ploidy))) + geom_line() + scale_color_discrete(name="Ploidy") + scale_linetype_discrete(name="Dominance Scalar", labels=c("0.1", "0.5", "0.9")) + xlim(0,300) + theme_bw() + xlab("Generation") + ylab("Allele Frequency")
```

## Fitness values as a function of frequency
```{r}
getFits = function(s, dom, ploidy){
  dom_coeffs = getDomCoefs(dom, ploidy)
  k = ploidy
  dat = foreach(i = 1:51, .combine=rbind) %do% {
    p = seq(0,1,1/50)[i]
    q = 1 - p
    het_fits = 1 + dom_coeffs * s
    fits = c(1, het_fits, 1 + s) / (1 + s)
    i = seq(0, k)
    freqs = choose(k, i) * (p ^ (k - i)) * (q ^ i)
    w.bar = sum(freqs * rev(fits))
    var.w = sum((freqs * (rev(fits) - w.bar) ^ 2) / length(freqs))
    data.frame('s' = s, 'dom' = dom, 'ploidy' = ploidy, 'freq' = p, 'w.bar' = w.bar, 'var.w' = var.w)
  }
  return(dat)
}

dat2 = foreach(i = 1:nrow(params), .combine=rbind) %do% {
  getFits(params[i,]$s, params[i,]$dom, params[i,]$ploidy)
}

ggplot(dat2, aes(x=freq, y=var.w, color = as.factor(ploidy), linetype=as.factor(dom))) + geom_line() + scale_color_discrete(name="Ploidy") + scale_linetype_discrete(name="Dominance", labels=c("0.1","0.5","0.9")) + xlab("Allele Frequency") + ylab("Variance in Fitness") + theme_bw()

ggplot(dat2, aes(x=freq, y=w.bar, color = as.factor(ploidy), linetype=as.factor(dom))) + geom_line() + scale_color_discrete(name="Ploidy") + scale_linetype_discrete(name="Dominance", labels=c("0.1","0.5","0.9")) + xlab("Allele Frequency") + ylab("Mean Fitness") + theme_bw()
```

## Variable key
N = Population size
s = selection coefficient; difference in fitness between alternative homozygotes
dom = dominance scalar; Run below chunk for meaning
recomb = per-base mutation rate
mu = per-base mutation rate.  Theta is ploidy x mu x N (x sequence length which is 1Mb)
sim_ploidy = ploidy used for mssel simulation.  Value used in calculating theta above
traj_ploidy = ploidy used to generate sweep trajectory.  see getTraj() function in PloidyHitch.R
sampGen = number of generations passed since selection ends (i.e. fixation of beneficial allele)
fuseGen = number of generations prior to start of sweep that the two populations fused.

## Summary of depth and breadth of peaks
Controlling for number of chromosomes in the population (i.e. Simulation Ploidy), higher ploidies have a shallower dip, due to the slowed rate of fixation of the sweeping allele. 
Similarly, the dips in diversity are narrower in higher ploidies for same reason as before. 
```{r}
MS2.a = ms2 %>% group_by(s, dom, recomb, N, mu, sim_ploidy, traj_ploidy, sampGen, fuseGen, rep) %>% summarize(depth = (mean(Pi.2,na.rm=T) - min(Pi.1)) / mean(Pi.2,na.rm=T))

MS2.a %>% filter(fuseGen == 1 & mu==1e-08 & N == 10000 & recomb=="1e-08" & s =="0.01") %>% ggplot(., aes(y=-(log(depth)), fill = as.factor(traj_ploidy), x = as.factor(sim_ploidy))) + geom_boxplot() + ylab(expression(paste("Dip Depth = min(", pi[1], ") / ",pi[2],sep = ""))) + facet_grid(~sampGen) + scale_y_log10() + scale_fill_discrete(name = "Trajectory\nPloidy") + xlab("Simulation Ploidy") + theme_bw()

MS2.b = ms2 %>% group_by(s, dom, recomb, N, mu, sim_ploidy, traj_ploidy, sampGen, fuseGen, rep, ints) %>% summarize(lessWind = ifelse(mean(Pi.1, na.rm=T) < (mean(Pi.2, na.rm=T) / 2),1,0))

MS2.c = MS2.b %>% group_by(s, dom, recomb, N, mu, sim_ploidy, traj_ploidy, sampGen, fuseGen, rep) %>% summarize(breadth = sum(lessWind) * 20000)

MS2.c %>% filter(fuseGen == 1 & mu==1e-08 & N == 10000 & recomb=="1e-08" & s =="0.01") %>% ggplot(., aes(y=breadth/1000, fill = as.factor(traj_ploidy), x = as.factor(sim_ploidy))) + geom_boxplot(notch=T) + ylab(expression(paste("Dip Breadth = Length (kb) of which ", pi[1], " < (",pi[2], " / 2)",sep = ""))) + facet_grid(~sampGen, labeller = label_both) + scale_fill_discrete(name = "Trajectory\nPloidy") + xlab("Simulation Ploidy") + theme_bw()

MS2.c %>% filter(fuseGen == 1 & mu==1e-08 & N == 10000 & recomb=="1e-08" & s =="0.01") %>% ggplot(., aes(y=breadth/1000, fill = as.factor(sampGen), x = as.factor(sim_ploidy))) + geom_boxplot(notch=T) + ylab(expression(paste("Dip Breadth = Length (1 kb) of which ", pi[1], " < (",pi[2], " / 2)",sep = ""))) + scale_y_log10() + scale_fill_discrete(name = "Generations prior\nto fixation") + xlab("Simulation Ploidy") + theme_bw()

# # Look for proportional difference in breadth/depth within a simul
# MS2.a %>% filter(fuseGen == 1 & mu==1e-08 & traj_ploidy == 4 & N == 10000 & recomb=="1e-08" & s =="0.01") %>% ggplot(., aes(y=depth, fill = as.factor(sampGen), x = as.factor(sim_ploidy))) + geom_boxplot(notch=T) + ylab(expression(paste("Dip Breadth = Length of which ", pi[1], " < (",pi[2], " / 2)",sep = "")))+ scale_fill_discrete(name = "Trajectory\nPloidy") + xlab("Simulation Ploidy") + theme_bw()
```

## Combined effects of factor of two and trajectory
Higher ploidies result in deeper and broader dips in diversity relative to diploids.  The effect of recombination rate is not striking.
```{r}
ms2 %>% filter(dom %in% c("0.1","0.9","0.5") & fuseGen == 1 & sampGen ==1 & mu==1e-08 & N == 10000 & sim_ploidy == traj_ploidy) %>% ggplot(., aes(y=Pi.1 * 100, color = as.factor(traj_ploidy), x = mid/100000)) + geom_smooth(method="loess", size =0.5, span=0.1, se=F) + scale_x_continuous(breaks=c(1,5,9)) + scale_y_continuous(breaks=c(0.05,0.15)) + xlab("Position (100 kb)") + ylab("Diversity (%)") + scale_color_discrete(name="Ploidy") + facet_grid(vars(recomb,s), vars(dom),labeller = label_both) + theme_bw()
```

## Trajectory effect
Here, we are keeping fixed the ploidy specified for the coalescent simulations (fixed at 8 in this example), and we are varying the ploidy under which the sweep trajectories are simulated.  For some parameter sets, higher ploidies and thus slower sweeps results in a visible difference in terms of the reduction of diversity at and around the beneficial allele.  The effect is more pronounced for weaker selection and for partial dominance/recessivity.  The breadth of the dips is also notably lesser for higher ploidies.  Thus, if chromosome number in a population is held constant, higher ploidies exhibit reduced signals associated with selective sweeps.
```{r}
ms2 %>% filter(dom %in% c("0.1","0.9","0.5") & N == 10000 & fuseGen == 1 & sampGen ==1 & mu==1e-08 & sim_ploidy == 8) %>% ggplot(., aes(y=Pi.1 * 100, color = as.factor(traj_ploidy), x = mid/100000)) + geom_smooth(method="loess", size =0.5, span=0.1, se=F) + scale_x_continuous(breaks=c(1,5,9)) + xlab("Position (100 kb)") + ylab("Diversity (%)") + scale_color_discrete(name="Trajectory\nPloidy") + facet_grid(vars(s, recomb), vars(dom),labeller = label_both, scales = "free_y") + theme_bw()
```

## Factor of 2 effect
This is the complement to the previous figure.  We use the same sweep trajectory, but we vary the ploidy specified for the coalescent simulation.  
```{r}

ms2 %>% filter(dom %in% c("0.1","0.9","0.5") & fuseGen == 1 & sampGen ==1 & mu==1e-08 & N == 10000 & traj_ploidy == 4) %>% ggplot(., aes(y=Pi.1 * 100, color = as.factor(sim_ploidy), x = mid/100000)) + geom_smooth(method="loess", size =0.5, span=0.1, se=F) + scale_x_continuous(breaks=c(1,5,9)) + scale_y_continuous(breaks=c(0.05,0.15)) + xlab("Position (100 kb)") + ylab("Diversity (%)") + scale_color_discrete(name="Simulation\nPloidy") + facet_grid(vars(recomb,s), vars(dom),labeller = label_both) + theme_bw()

```

## Effect of sample generation on Taj D.  Polyploids recover new mutations more quickly, but they take longer to drift to higher frequencies.  ThetaW goes up but ThetaPi lags...
```{r}
ms2 %>% filter(dom %in% c("0.1","0.9","0.5") & s==0.01 & fuseGen == 1 & mu==1e-08 & N == 10000 & sim_ploidy == traj_ploidy & recomb == 1e-08) %>% ggplot(., aes(y=Tajima.D_1, color = as.factor(sim_ploidy), x = mid/100000)) + geom_smooth(method="loess", size =0.5, span=0.1, se=F) + scale_x_continuous(breaks=c(1,5,9)) + xlab("Position (100 kb)") + ylab("Tajima's D") + scale_color_discrete(name="Ploidy") + facet_grid(vars(sampGen), vars(dom),labeller = label_both) + theme_bw()

ms2 %>% filter(dom == "0.5" & s==0.01 & fuseGen == 1 & mu==1e-08 & N == 10000 & recomb == 1e-08) %>% ggplot(., aes(y=Tajima.D_1, color = as.factor(sim_ploidy), x = mid/100000)) + geom_smooth(method="loess", size =0.5, span=0.1, se=F) + scale_x_continuous(breaks=c(1,5,9)) + xlab("Position (100 kb)") + ylab("Tajima's D") + scale_color_discrete(name="Ploidy") + facet_grid(vars(sampGen), vars(traj_ploidy),labeller = label_both) + theme_bw()
```

## Effect of sample generation on Fst.  
Neutral divergence in allele frequencies increases Fst more quickly in lower ploidies, thus erasing 'outlier' signal due to sweep
```{r}
ms2 %>% filter(dom %in% c("0.1","0.9","0.5") & s==0.01 & fuseGen == 1 & mu==1e-08 & N == 10000 & sim_ploidy == traj_ploidy & recomb == 1e-08) %>% ggplot(., aes(y=fst, color = as.factor(sim_ploidy), x = mid/100000)) + geom_smooth(method="loess", size =0.5, span=0.1, se=F) + scale_x_continuous(breaks=c(1,5,9)) + xlab("Position (100 kb)") + ylab("Fst") + scale_color_discrete(name="Ploidy") + facet_grid(vars(sampGen), vars(dom),labeller = label_both) + theme_bw()
```

##Analysis of rehh calculations
###Import Data
```{r}
traj.ehh = read.csv("~/Documents/Research/PloidySim/rehh/rehh_calcs_sp2_se0.01.csv")
fact.ehh = read.csv("~/Documents/Research/PloidySim/rehh/rehh_calcs_do0.5_fG1_DS.csv")
```

###Trajectory effect on xp-ehh (colors) 
```{r}
traj.ehh %>% filter(fuseGen=="fG1" & sampGen=="sG1" & recomb == "re1e-08" & dom=="do0.5" & N=="NN10000") %>% ggplot(.,aes(x=POSITION / 100000,y=xpehh.p,color=traj_ploidy)) + geom_smooth(method="loess", span=0.1, se = F) + scale_color_discrete(name="Trajectory\nPloidy", breaks=c("tp2","tp4","tp8"), labels=c("2","4","8")) + xlab("Position (100 kb)") + ylab(expression("-log"[10]*"(p-value)")) + theme_bw()

# fact.ehh %>% filter(sim_ploidy=="sp8" & fuseGen=="fG1" & sampGen=="sG1" & recomb == "re1e-08" & dom=="do0.5" & N=="NN10000") %>% ggplot(.,aes(x=POSITION / 100000,y=xpehh.p,color=traj_ploidy)) + geom_smooth(method="loess", span=0.1, se = F) + scale_color_discrete(name="Trajectory\nPloidy", breaks=c("tp2","tp4","tp8"), labels=c("2","4","8")) + xlab("Position (100 kb)") + ylab(expression("-log"[10]*"(p-value)")) + theme_bw()
```

###Factor of two effect on xp-ehh for different trajectory ploidies
That is, trajectory ploidy is constant within each panel, while simulation ploidy is varied.  Higher diversity in higher ploidies produces stronger signal due to greater deviation from baseline genomic diversity.

Something strange here though...why is p value always highest when tp=sp?
Maybe look at density of NA along chromosomes?
```{r}
# fact.ehh %>% filter(traj_ploidy == "tp4" & fuseGen=="fG1" & dom=="do0.5" & N=="NN10000" & recomb=="re1e-08" & sampGen=="sG1") %>% ggplot(.,aes(x=cut_interval(x=POSITION, n=50),y=xpehh.p,fill=sim_ploidy)) + geom_boxplot() + facet_grid(cols = vars(sampGen))

fact.ehh %>% filter(s == "se0.01" & fuseGen=="fG1" & recomb == "re1e-08" & dom=="do0.5" & N=="NN10000") %>% ggplot(.,aes(x=POSITION / 100000, y=nocalc, color=sim_ploidy)) + geom_smooth(method="loess", span=0.1, se = F) + scale_color_discrete(name="Simulation\nPloidy", breaks=c("sp2","sp4","sp8"), labels=c("2","4","8"))  + facet_grid(cols = vars(traj_ploidy), rows = vars(sampGen))  + xlab("Position (100 kb)") + ylab(expression("-log"[10]*"(p-value)")) + theme_bw()

```

###Effect of sample generation for each ploidy.  Does signal decay faster in higher or lower ploidies?
```{r}
fact.ehh %>% filter(s == "se0.01" & fuseGen=="fG1" & dom=="do0.5" & N=="NN10000" & recomb=="re1e-08") %>% ggplot(.,aes(x=POSITION,y=xpehh.p,color=sim_ploidy)) + geom_smooth() + facet_grid(rows = vars(traj_ploidy), cols = vars(sampGen))
```

# Less interesting analyses

## Effect of sample generation
Sample generation is the number of generations following a sweep at which sampling occurs
```{r} 
ms2 %>% filter(dom %in% c("0.1","0.9","0.5") & s==0.01 & fuseGen == 1 & mu==1e-08 & N == 10000 & recomb == 5e-08 & sim_ploidy == traj_ploidy) %>% ggplot(., aes(y=Pi.1 * 100, color = as.factor(sim_ploidy), x = mid/100000)) + geom_smooth(method="loess", size =0.5, span=0.1, se=F) + scale_x_continuous(breaks=c(1,5,9)) + scale_y_continuous(breaks=c(0.05,0.15)) + xlab("Position (100 kb)") + ylab("Diversity (%)") + scale_color_discrete(name="Simulation\nPloidy") + facet_grid(vars(sampGen), vars(dom),labeller = label_both) + theme_bw()
```

## Effect of population size
Nothing really interesting here...Smaller population size seems noisier as expected
```{r}
ms2 %>% filter(dom %in% c("0.1","0.9","0.5") & recomb=="1e-08" & fuseGen == 1 & sampGen ==1 & mu==1e-08 & sim_ploidy == traj_ploidy) %>% ggplot(., aes(y=Pi.1 * 100, color = as.factor(traj_ploidy), x = mid/100000)) + geom_smooth(method="loess", size =0.5, span=0.1, se=F) + scale_x_continuous(breaks=c(1,5,9)) + xlab("Position (100 kb)") + ylab("Diversity (%)") + scale_color_discrete(name="Ploidy") + facet_grid(vars(s, N), vars(dom),labeller = label_both, scales = "free_y") + theme_bw()
```