### 2019-10-12 analysis/PaperFigures.Rmd
resubmission changes

---
title: "Paper_Figures"
author: "Patrick Monnahan"
date: "8/27/2019"
output: html_document
---

## Load packages, import data, and define functions
```{r setup, message=F}
knitr::opts_chunk$set(echo = TRUE)
library(assertthat)
library(ggplot2)
library(dplyr)
library(tidyr)
library(PopGenome)
library(stringr)
library(data.table)
library(magrittr)
library(wrapr)
library(foreach)
library(purrr)
library(IDPmisc)
library(MESS)
library(wesanderson)

getFits = function(s, dom, ploidy){
  dom_coeffs = getDomCoefs(dom, ploidy)
  k = ploidy
  dat = foreach(i = 1:51, .combine=rbind) %do% {
    p = seq(0,1,1/50)[i]
    q = 1 - p
    het_fits = 1 + dom_coeffs * s
    fits = c(1, het_fits, 1 + s) / (1 + s)
    i = seq(0, k)
    freqs = choose(k, i) * (p ^ (k - i)) * (q ^ i)
    w.bar = sum(freqs * rev(fits))
    var.w = sum((freqs * (rev(fits) - w.bar) ^ 2) / length(freqs))
    data.frame('s' = s, 'dom' = dom, 'ploidy' = ploidy, 'freq' = p, 'w.bar' = w.bar, 'var.w' = var.w)
  }
  return(dat)
}

dipTraj=function(start_freq,s,h,stop_freq){
  h=1-h
  p=start_freq
  traj = c(p)
  fits = c(1.0-s, 1.0 - h*s, 1.0)
  while(p < stop_freq){
    g_freqs = c((1-p)^2, 2*p*(1 - p), p^2)
    tot_fit = sum(fits * g_freqs)
    p_fit = (g_freqs[3] * fits[3]) + 0.5*(g_freqs[2] * fits[2])
    p_prime = p_fit/tot_fit
    traj = c(traj,p_prime)
    p = p_prime
  }
  return(traj)
}

tetTraj=function(start_freq,s,h1,h2,h3,stop_freq){
  h1 = 1-h1
  h2 = 1-h2
  h3 = 1-h3
  p=start_freq
  traj = c(p)
  fits = c(1.0-s, 1.0 - h1*s, 1.0 - h2*s, 1.0 - h3*s, 1.0)
  while(p < stop_freq){
    g_freqs = c((1-p)^4, 4*p*(1 - p)^3, 6*p^2*(1 - p)^2, 4*p^3*(1 - p), p^4)
    tot_fit = sum(fits * g_freqs)
    p_fit = (g_freqs[5] * fits[5]) + 0.75*(g_freqs[4] * fits[4]) + 0.5*(g_freqs[3] * fits[3]) + 0.25*(g_freqs[2] * fits[2])
    p_prime = p_fit/tot_fit
    traj = c(traj,p_prime)
    p = p_prime
  }
  return(traj)
}

getDomCoefs = function(dominance, ploidy = 100){
  if (dominance == 0){
    coeffs = rep(0, ploidy - 1)
  }
  else if (dominance == 1){
    coeffs = rep(1, ploidy - 1)
  }
  else {
    dom = (1 - dominance) - 0.5
    dom = abs((dom * 10 + sign(dom)) ^ sign(dom))
    coeffs = (seq(1, ploidy - 1) / ploidy) ^ dom
  }
  return(coeffs)
}

getStat = function(dat, stat, offset = 25000){
  dat = dat %>% filter(mid > 500000 - offset & mid < 500000 + offset)
  val = max(abs(unlist(dat[,stat])), na.rm = T)
  return(val)
  }

getTraj = function(s, dom, ploidy, start_freq, N = -9, end_freq = 0.99, maxGens = 9999, maxTries = 10){
  dom_coeffs = getDomCoefs(dom, ploidy)
  attempts = 0
  p = start_freq
  k = ploidy
  gen = 0
  while (attempts <= maxTries & p < end_freq & gen <= maxGens){
    df = data.frame("s" = s, "dom" = dom, "gen" = 0, "freq" = start_freq, "dp1" = 0, "w.bar" = 0, "ploidy" = ploidy)
    p = start_freq
    dp1 = 0
    gen = 0
    het_fits = 1 + dom_coeffs * s
    fits = c(1, het_fits, 1 + s) / (1 + s)
    while (p < end_freq & p > 0 & gen <= maxGens){
      q = 1 - p
      i = seq(0, k)
      Num = sum(choose(k, i) * ((k - i) / k) * (p ^ (k - i)) * (q ^ i) * rev(fits))
      Den = sum(choose(k, i) * (p ^ (k - i)) * (q ^ i) * rev(fits))
      p_prime = Num / Den
      if (N != -9){
        p_prime = sum(rbinom(N, k, p_prime)) / (k * N) 
      }
      df = rbind(df, c(s, dom, gen, p, dp1, Den, k))
      dp1 = p_prime - p
      p = p_prime
      gen = gen + 1
    }
    df = rbind(df, c(s, dom, gen, p, dp1, Den, k))
    attempts = attempts + 1
  }
  if (attempts > maxTries & nrow(df) <= maxGens){
    print(paste("Beneficial allele was lost due to drift for", maxTries, "consecutive attempts"))
  }
  if (nrow(df) > maxGens){
    print(paste("Beneficial allele did not fix before the maximum number of generations (", maxGens, ")."))
  }
  return(df[-1,])
}

getDims = function(dat, y.metric = "fst", offset = 100000){
  # dat = data.frame("mid" = mid, "fst" = febugst)
  peak = tryCatch({
    auc(dat$mid, unlist(abs(dat[,y.metric])), from = 500000 - offset, to = 500000 + offset)}
    , error = function(e){
      return(-9)
    })
  return(peak)
}

getUnd = function(dat, y.metric = "fst", offset = 100000){
  # dat = data.frame("mid" = mid, "fst" = febugst)
  dat = dat %>% filter(mid > 500000 - offset & mid < 500000 + offset)
  dat["stat"] = dat[, y.metric]
  val1 = nrow(dat[dat$stat == "-Inf",])
  val2 = nrow(dat[is.na(dat$stat),])
  return(val1+val2)
}

getUnd.h = function(dat, y.metric = "fst", offset = 100000){
  # dat = data.frame("mid" = mid, "fst" = febugst)
  dat = dat %>% filter(POSITION > 500000 - offset & POSITION < 500000 + offset)
  dat["stat"] = dat[, y.metric]
  val1 = nrow(dat[dat$stat == "-Inf",])
  val2 = nrow(dat[is.na(dat$stat),])
  return(val1+val2)
}

getStat = function(dat, stat, offset = 25000){
  dat = dat %>% filter(mid > 500000 - offset & mid < 500000 + offset)
  val = max(abs(unlist(dat[,stat])), na.rm = T)
  return(val)
  }

wrapEHH = function(files){
  all.sum = data.frame()
  for(i in 1:length(files)){
    print(files[i])
    dat = read.csv(files[i])
    dat.sum = dat %>% mutate(traj_ploidy = as.numeric(str_replace(traj_ploidy, "tp","")), sim_ploidy = as.numeric(str_replace(sim_ploidy, "sp","")), dom = as.numeric(str_replace(dom, "do","")), s = as.numeric(str_replace(s, "se","")), N = as.numeric(str_replace(N, "NN","")), mu = as.numeric(str_replace(mu, "mu","")), recomb = as.numeric(str_replace(recomb, "re","")), sampGen = as.numeric(str_replace(sampGen, "sG","")), fuseGen = as.numeric(str_replace(fuseGen, "fG","")), rep = as.numeric(str_replace(rep, "rep", ""))) %>% filter(!is.na(xpehh.p)) %>% mutate(mid = POSITION) %>% nest(-c(sim_ploidy, dom, rep)) %>% mutate(size.xp = map(data, ~getDims(., y.metric = "XPEHH")), size.xp.p = map(data, ~getDims(., y.metric = "xpehh.p")), med.xp = map(data, ~getStat(., "XPEHH")), med.xpp = map(data, ~log(getStat(., "xpehh.p"))), size.ihs = map(data, ~getDims(., y.metric = "IHS")), size.ihs.p = map(data, ~getDims(., y.metric = "ihs.p")), med.ihs = map(data, ~getStat(., "IHS")), med.ihs.p = map(data, ~log(getStat(., "ihs.p")))) %>% select(-data) %>% unnest()
    all.sum = rbind(all.sum, dat.sum)
    remove(dat)
  }
  return(all.sum)
}

CV <- function(data){
  data = data[is.finite(data)]
  return((sd(data, na.rm=T)/(mean(data, na.rm=T)))*100)
  }

lCV <- function(data, N = 10000, n = 100){
    aa = foreach(i = 1:N, .combine = rbind) %do% {
      ndat = sample(data, n)
      sd(ndat, na.rm=T)/(mean(ndat, na.rm=T))*100}
    return(quantile(aa,0.05, na.rm=T))
}

uCV <- function(data, N = 10000, n = 100){
    aa = foreach(i = 1:N, .combine = rbind) %do% {
      ndat = sample(data, n)
      sd(ndat, na.rm=T)/(mean(ndat, na.rm=T))*100}
    return(quantile(aa,0.95, na.rm=T))
}

#Load data
scaleFUN <- function(x) sprintf("%.1f", x)

# pardir = "/Users/..." #Directory where data files are located
pardir = "~/Documents/Research/PloidySim/data/"
ms = read.table(paste(pardir, "mssel_out_10-4-19.txt", sep = ""), head = T)
ms %<>% mutate(ints = as.numeric(as.character(cut(bp.start, breaks = 50, labels = seq(1,50))))) # label (bin) windows into larger intervals

ms2 = ms %>% filter(DS != "-9")
ms %<>% filter(DS == "-9")

ds.ehh.files = c(paste(pardir,"rehh_DS.csv", sep = ""))
nds.ehh.files = c(paste(pardir,"rehh_nDS.csv", sep = ""))
traj=read.table(paste(pardir, "Ploidy_Forward_Sim_results.txt", sep = ""),head=T, sep=",")

rehh = read.csv(paste(pardir, "rehh_calcs_do0.5_fG1_nDS.csv", sep=""))

ds.ehh = wrapEHH(ds.ehh.files)
nds.ehh = wrapEHH(nds.ehh.files)
ds.ehh %<>% mutate(DS="yes")
nds.ehh %<>% mutate(DS="no")
ehh = rbind(nds.ehh, ds.ehh)
remove(ds.ehh)
remove(nds.ehh)
```

## Sweep trajectories; Figure 1A.

These plots show a primary motivating result for this study, in that higher ploidies show increased fixation times of sweeping beneficial alleles regardless of the dominance scenario being evaluated.

```{r, echo=T,message=F}
#Function creates trajectories

# Param set to create trajectories for
params = data.frame("s" = rep(0.1,9), "dom" = c(rep(1,3), rep(0.5,3), rep(0,3)), "ploidy" = rep(c(2,4,8),3), "start" = rep(0.05, 9))

# Create trajectories
dat1 = foreach(i = 1:nrow(params), .combine=rbind) %do% {
  getTraj(params[i,]$s, params[i,]$dom, params[i,]$ploidy, params[i,]$start, maxGens = 99999, maxTries = 1)
}

dat1$dom = as.factor(dat1$dom)
dat1$dom = factor(dat1$dom, levels(dat1$dom)[c(3,2,1)])

#Figure 1A left
dat1 %>% filter(ploidy %in% c(2, 4, 8) & dom %in% c(0, 0.5, 1)) %>% ggplot(., aes(x=gen, y=freq, linetype=as.factor(dom), color=as.factor(ploidy))) + geom_line(size = 1.2) + scale_color_manual(name = "Ploidy", values = c("#F8766D", "#00BA38", "#619CFF")) + scale_linetype_manual(name = "Dominance", labels=c("Dominant", "Additive", "Recessive"), values = c("dashed", "solid", "dotted")) + xlim(0,500) + theme_bw() + xlab("Generation") + ylab("Allele Frequency") + theme(axis.text.y=element_text(size=16), axis.title.y=element_text(size=18),axis.text.x=element_text(size=16),axis.title.x=element_text(size=18),legend.text = element_text(size=16), legend.title = element_text(size=18))

dat2 = foreach(i = 1:nrow(params), .combine=rbind) %do% {
  getFits(params[i,]$s, params[i,]$dom, params[i,]$ploidy)
}

dat2$dom = as.factor(dat2$dom)
dat2$dom = factor(dat2$dom, levels(dat2$dom)[c(3,2,1)])

#Figure 1A right
ggplot(dat2, aes(x=freq, y=var.w, color = as.factor(ploidy), linetype=as.factor(dom))) + geom_line(size = 0.9) + scale_color_discrete(name="Ploidy") + scale_linetype_manual(name="Dominance", labels=c("Dominant","Additive","Recessive"), values = c("dashed", "solid", "dotted")) + xlab("Allele Frequency") + ylab("Variance in Fitness\n") + theme_bw() + theme(axis.text.y=element_blank(), axis.title.y=element_text(size=18),axis.text.x=element_text(size=16),axis.title.x=element_text(size=18),legend.text = element_text(size=16), legend.title = element_text(size=18))

ggplot(dat2, aes(x=freq, y=w.bar, color = as.factor(ploidy), linetype=as.factor(dom))) + geom_line() + scale_color_discrete(name="Ploidy") + scale_linetype_discrete(name="Dominance", labels=c("0.1","0.5","0.9")) + xlab("Allele Frequency") + ylab("Mean Fitness") + theme_bw()
```

## Dominance; Figure 1B

This plot shows the dominance value of each genotype (given by within individual allele frequency) as a function of the dominance scalar.  The purpose was to write a function that would take a single dominance value that could be specified for any ploidy and return a comparable set of dominance coefficients for heterozygous genotypes.

```{r, echo=T,message=F}

doms = c(seq(0,1, 0.1))
Doms = as.data.frame(sapply(doms, getDomCoefs))
Doms = rbind(c(0,0,0,0,0,0,0), Doms, c(1,1,1,1,1,1,1))
colnames(Doms) = doms
Doms = cbind(Doms, data.frame("i/k"=seq(0,100) / 100))
mDoms = melt(Doms, id.var = "i.k")

mDoms$variable = factor(mDoms$variable, levels(mDoms$variable)[c(rev(seq(1,11)))])

mDoms %>% filter(variable %in% c("0.9", "0.6", "0.5", "0.4", "0.1")) %>% ggplot(., aes(x = i.k, y = value, color = variable)) + geom_line(size = 1.2) + scale_color_discrete(name = "Dominance\nScalar", breaks = c("0.9", "0.6", "0.5", "0.4", "0.1"), labels = c("0.4", "0.1", "0", "-0.1", "-0.4")) + theme_bw() + xlab("Individual allele frequency of beneficial allele") + ylab("Dominance Coefficient") + theme(axis.text.y=element_text(size=16), axis.title.y=element_text(size=18),axis.text.x=element_text(size=16),axis.title.x=element_text(size=18),legend.text = element_text(size=16), legend.title = element_text(size=18))
```

## Stochasticity in recessive trajectories; Figure 1C
```{r}
bbR=data.frame(traj=dipTraj(0.05,0.1,0,0.99))
bbR$gen=seq.int(nrow(bbR))
aaR=data.frame(traj=tetTraj(0.05,0.1,0,0,0,0.99))
aaR$gen=seq.int(nrow(aaR))

traj %>% filter(N %in% c(1000,10000) & s == 0.01 & dom == "recessive") %>% ggplot()+geom_line(data=traj[traj$dom == "recessive" & traj$s==0.1 & traj$ploidy==4 & traj$N %in% c(1000,10000),],aes(y=freq,x=gen,group=as.factor(rep)),color="blue",alpha=0.5)+geom_line(data=traj[traj$dom == "recessive" & traj$s==0.1 & traj$ploidy==2 & traj$N %in% c(1000,10000),],aes(y=freq,x=gen,group=as.factor(rep)),color="red",alpha=0.5)+geom_line(data=bbR,aes(y=traj,x=gen), linetype="dashed")+geom_line(data=aaR,aes(y=traj,x=gen))+facet_grid(~N,scales="free_x")+xlab("Generation")+ylab("Frequency")+theme(legend.title = element_text("Ploidy"))+scale_x_log10()
```

## Effect of population size on fixation time; Figure 1D
```{r, echo=T,message=F}
#Stochastic effect of population size on
params2 = data.frame("s" = rep(0.1,12), "dom" = c(rep(0.1,4), rep(0.5,4), rep(0.9,4)), "ploidy" = rep(c(2,4), 6), "start" = rep(0.05, 12), "N" = rep(c(1000, 1000, 10000, 10000), 3))
params2 %<>% dplyr::slice(rep(row_number(), 1000))
# Create trajectories
dat3 = foreach(i = 1:nrow(params2), .combine=rbind) %do% {
  print(params2[i,])
  dat = getTraj(params2[i,]$s, params2[i,]$dom, params2[i,]$ploidy, params2[i,]$start, N = params2[i,]$N, maxGens = 9999999, maxTries = 9999999)
  dat["N"] = params2[i,]$N
  dat[nrow(dat),]
}

dat3 %>%  mutate(Dom = factor(case_when(dom == 0.1 ~ "Partial\nRecessivity (H = -0.4)", dom == 0.5 ~ "Additivity (H = 0)", dom == 0.9 ~ "Partial\nDominance (H = 0.4)"), levels = c("Partial\nRecessivity (H = -0.4)", "Additivity (H = 0)", "Partial\nDominance (H = 0.4)"))) %>% ggplot(aes(y = log(gen), fill = as.factor(N), x = as.factor(ploidy))) + geom_boxplot(outlier.size = 0.5) + facet_wrap(~Dom) + ylab("Log(Time to Fixation)") + xlab("Ploidy") + scale_fill_discrete(name = "N")+ theme_bw() + theme(axis.text.y=element_text(size=14), axis.title.y=element_text(size=16),axis.text.x=element_text(size=14),axis.title.x=element_text(size=16),legend.text = element_text(size=14), legend.title = element_text(size=16), strip.text = element_text(size=12))

```

---

# mssel analysis

### Variable key
* N = Population size

* s = selection coefficient; difference in fitness between alternative homozygotes

* dom = dominance scalar; a single value that determines the dominance coefficients of heterozygous genotypes for an arbitrary ploidy

* recomb = per-base recombination rate

* mu = per-base mutation rate.  Theta is ploidy x mu x N (x sequence length which is 1Mb)

* sim_ploidy = **Simulation Ploidy**; ploidy used for mssel simulation.  Value used in calculating theta above

* traj_ploidy = **Trajectory Ploidy**; ploidy used to generate sweep trajectory.  see getTraj() function in PloidyHitch.R

* sampGen = sampling generation; number of generations passed since selection ends (i.e. fixation of beneficial allele)

* fuseGen = fuse generation; number of generations prior to start of sweep that the two populations fused.

**Simulation Ploidy** and **Trajectory Ploidy** are the two focal variables.  The former refers to the ploidy value  %<>%  %>%used to scale population parameters when running mssel.  For example, theta = 2(N)(mu)(sim_ploidy).  The latter refers to the ploidy used for simulating the sweep trajectory, which is subsequently passed to mssel.

## Figure 2A.  
```{r, echo=T,message=F}
#Figure 2A Smoothed diversity versus position
ms %>% filter(traj_ploidy==sim_ploidy & sampGen == "1" & dom == "0.5" & N == 10000) %>% ggplot(., aes(y=Pi.1 * 100, color = as.factor(sim_ploidy), x = mid/100000, linetype = as.factor(s))) + geom_smooth(method="loess", size = 0.9, span=0.1, se=F) + scale_x_continuous(breaks=c(1,5,9), limits = c(0.2,9.8)) + xlab("Position (100 kb)") + ylab("Diversity (%)") + scale_linetype_discrete(name="Selection\nCoefficient") + scale_color_discrete(name="Ploidy") + theme_bw() + theme(axis.text.y=element_text(size=16), axis.title.y=element_text(size=18),axis.text.x=element_text(size=16),axis.title.x=element_text(size=18),legend.text = element_text(size=16), legend.title = element_text(size=18))
```

#Figure 2B
```{r}
MS2.a = ms %>% group_by(s, dom, recomb, N, mu, sim_ploidy, traj_ploidy, sampGen, fuseGen, rep, driftGen) %>% summarize(depth.pi = mean(Pi.2,na.rm=T) - min(Pi.1), min.pi = min(Pi.1, na.rm=T), max.taj = abs(min(Tajima.D_1, na.rm = T)), height.fst = max(fst, na.rm = T) - mean(fst, na.rm = T))

MS2.b = ms %>% group_by(s, dom, recomb, N, mu, sim_ploidy, traj_ploidy, sampGen, fuseGen, driftGen, rep, ints) %>% summarize(lessWind.pi = ifelse(mean(Pi.1, na.rm=T) < (mean(Pi.2, na.rm=T) / 2),1,0), lessWind.taj = ifelse(mean(Tajima.D_1, na.rm=T) < -1, 1, 0))

MS2.c = MS2.b %>% group_by(s, dom, recomb, N, mu, sim_ploidy, traj_ploidy, sampGen, fuseGen, driftGen, rep) %>% summarize(breadth.pi = sum(lessWind.pi) * 20000, breadth.taj = sum(lessWind.taj) * 20000)

MS2.f = merge(MS2.a, MS2.c, by = c("s", "dom", "recomb", "N", "mu", "sim_ploidy", "traj_ploidy", "sampGen", "fuseGen", "driftGen", "rep"))

MS2.f %<>% mutate(same_ploidy = ifelse(sim_ploidy==traj_ploidy, "yellow", "black"))

#disentangling
# MS2.f %>% filter(mu==1e-08 & N == 10000 & s =="0.01" & driftGen %in% c(0,1) & sampGen == 1 & dom %in% c(0.5)) %>% mutate(Height = (log(min.pi, 10)), Breadth = breadth.pi/10000, Magnitude = depth.pi) %>% select(sim_ploidy, traj_ploidy, Magnitude, Breadth, same_ploidy) %>% melt(., id.var = c("sim_ploidy", "traj_ploidy", "same_ploidy")) %>% ggplot(., aes(y = value, fill = as.factor(traj_ploidy), x = as.factor(sim_ploidy), color = same_ploidy)) + geom_boxplot(outlier.shape = NA) + scale_fill_manual(name = "Trajectory Ploidy", values = wes_palette("Moonrise2")) + scale_x_discrete(name = expression(paste(theta)), labels = c(expression(paste("4x10", "^-4")), expression(paste("8x10", "^-4")), expression(paste("16x10", "^-4")))) + scale_color_manual(guide=F, values = c("black", "red")) + facet_wrap(~variable, scales = "free_y") + ylab("") + xlab("\n\n") + theme_bw() + theme(axis.text.y=element_text(size = 16), axis.title.y=element_text(size=18), axis.text.x=element_text(size=16), axis.title.x=element_text(size=18), legend.text = element_text(size=16), legend.title = element_text(size=18), strip.text = element_text(size=16))

MS2.f %>% filter(mu==1e-08 & N == 10000 & s =="0.01" & driftGen %in% c(0,1) & sampGen == 1 & dom %in% c(0.5)) %>% mutate(Height = (log(min.pi, 10)), Breadth = breadth.pi/100000, Magnitude = depth.pi * 100) %>% select(sim_ploidy, traj_ploidy, Magnitude, Breadth, same_ploidy) %>% melt(., id.var = c("sim_ploidy", "traj_ploidy", "same_ploidy")) %>% ggplot(., aes(y = value, fill = as.factor(traj_ploidy), x = as.factor(sim_ploidy))) + geom_boxplot(outlier.shape = NA) + scale_fill_manual(name = "Sweep\nTrajectory", values = wes_palette("Moonrise2"), labels = c("Diploid", "Tetraploid", "Octaploid")) + scale_x_discrete(name = expression(paste(theta," = ", rho)), labels = c(expression(paste("4x10"^"-4")), expression(paste("8x10"^"-4")), expression(paste("16x10"^"-4")))) + scale_color_manual(guide=F, values = c("black", "red")) + facet_wrap(~variable, scales = "free_y") + ylab("") + xlab("\n\n") + theme_bw() + theme(axis.text.y=element_text(size = 14), axis.title.y=element_text(size=18), axis.text.x=element_text(size=14), axis.title.x=element_text(size=22), legend.text = element_text(size=12), legend.title = element_text(size=14), strip.text = element_text(size=14))
```

#Figure 2C
```{r}
ff = rehh %>% filter(!is.na(xpehh.p) & sampGen!=1000) %>% mutate(mid = POSITION) %>% nest(-c(s, dom, recomb, N, mu, sim_ploidy, traj_ploidy, sampGen, fuseGen, rep)) %>% mutate(size.xp = map(data, ~getDims(., y.metric = "XPEHH")), size.xp.p = map(data, ~getDims(., y.metric = "xpehh.p")), med.xp = map(data, ~getStat(., "XPEHH")), med.xpp = map(data, ~log(getStat(., "xpehh.p"))), size.ihs = map(data, ~getDims(., y.metric = "IHS")), size.ihs.p = map(data, ~getDims(., y.metric = "ihs.p")), med.ihs = map(data, ~getStat(., "IHS")), med.ihs.p = map(data, ~log(getStat(., "ihs.p")))) %>% select(-c(data)) %>% unnest()

ff %<>% mutate(traj_ploidy = as.numeric(str_replace(traj_ploidy, "tp","")), sim_ploidy = as.numeric(str_replace(sim_ploidy, "sp","")), dom = as.numeric(str_replace(dom, "do","")), s = as.numeric(str_replace(s, "se","")), N = as.numeric(str_replace(N, "NN","")), mu = as.numeric(str_replace(mu, "mu","")), recomb = as.numeric(str_replace(recomb, "re","")), sampGen = as.numeric(str_replace(sampGen, "sG","")), fuseGen = as.numeric(str_replace(fuseGen, "fG","")), rep = as.numeric(str_replace(rep, "rep", "")))

ee = ms %>% filter(sampGen==1 & driftGen == 0 & fuseGen %in% c(1,1000,10000) & dom %in% c(0.1, 0.5, 0.9)) %>% group_by(s, dom, recomb, N, mu, sim_ploidy, traj_ploidy, sampGen, fuseGen, rep, driftGen) %>% mutate(Div = mean(Pi.2) - Pi.1, Tajima.D_1 = abs(Tajima.D_1)) %>% ungroup() %>% nest(-c(s, dom, recomb, N, mu, sim_ploidy, traj_ploidy, sampGen, fuseGen, rep, driftGen)) %>% mutate(size.div = map(data, ~getDims(., y.metric = "Div")), size.taj = map(data, ~getDims(., y.metric = "Tajima.D_1")), size.h = map(data, ~getDims(., y.metric = "Fay.Wu.H_1")), size.e = map(data, ~getDims(., y.metric = "Zeng.E_1")), size.fst = map(data, ~getDims(., y.metric = "fst")), size.dxy = map(data, ~getDims(., y.metric = "dxy.1.2")), med.d = map(data, ~getStat(., "Tajima.D_1")), med.e = map(data, ~getStat(., "Zeng.E_1")), med.h = map(data, ~getStat(., "Fay.Wu.H_1")), med.fst = map(data, ~getStat(., "fst")), med.dxy = map(data, ~getStat(., "dxy.1.2"))) %>% select(-data) %>% unnest()

kk = merge(ee, ff, by = c("traj_ploidy", "sim_ploidy", "dom", "s", "N", "mu", "recomb", "sampGen", "fuseGen", "rep"), all.x = T)

kk.points = kk %>% filter(dom %in% c("0.1", "0.5", "0.9") & traj_ploidy == sim_ploidy & s == "0.01" & recomb == "1e-08" & N=="1000" & driftGen == 0 & sampGen == 1)  %>% mutate("TajD" = size.taj, "FayWuH" = size.h, "ZengsE" = size.e, "Fst" = size.fst, "XPEHH" = size.xp, "Diversity" = size.div * 10) %>% select(sim_ploidy, dom, TajD, FayWuH, ZengsE, Fst, XPEHH, Diversity) %>% melt(., id.var = c("sim_ploidy", "dom")) %>% mutate("PeakSize" = variable) %>% filter(value != -9) %>% group_by(sim_ploidy, dom, variable) %>% summarize(median = median(value)) %>% mutate(PeakSize = variable)

#Just diversity for panel
kk %>% filter(dom %in% c(0.1, 0.5, 0.9) & traj_ploidy == sim_ploidy & s == "0.1" & N=="10000" & driftGen == 0 & sampGen == 1 & recomb == 1e-08 & fuseGen == 1) %>% mutate("Diversity" = size.div, same_ploidy = ifelse(sim_ploidy == traj_ploidy, 1, 0)) %>% select(sim_ploidy, dom, Diversity) %>% melt(., id.var = c("sim_ploidy", "dom")) %>% mutate("PeakSize" = variable) %>% filter(value != -9) %>% ggplot(., aes(y = value, fill = as.factor(dom), x = as.factor(sim_ploidy))) + geom_boxplot(outlier.shape = NA) + geom_point(data = kk.points[kk.points$variable == "Diversity",], aes(y = median, x = as.factor(sim_ploidy)), position = position_dodge(width = 0.8), size = 2, shape = 23, color="cyan") + scale_fill_manual(name = "Dominance\nScalar", values = wes_palette("Royal1"), labels = c("-0.4 (Recessive)", "\n0 (Additive)\n", "0.4 (Dominant)")) + ylab("Peak Area") + xlab("Ploidy") + theme_bw() + theme(axis.text.y=element_text(size = 12), axis.title.y=element_text(size=14), axis.text.x=element_text(size=12), axis.title.x=element_text(size=14), legend.text = element_text(size=12), legend.title = element_text(size=14), strip.text = element_text(size=16))

```

#Figure 2D
```{r}
dd = ms %>% filter(sim_ploidy == traj_ploidy & fuseGen==1 & s == "0.01" & recomb == "1e-08" & N=="10000" & driftGen == 0 & sampGen %in% c(1,1000,10000)) %>% group_by(s, dom, recomb, N, mu, sim_ploidy, traj_ploidy, sampGen, fuseGen, rep, driftGen) %>% mutate(Fst = scale(fst)) %>% ungroup() %>% nest(-c(s, dom, recomb, N, mu, sim_ploidy, traj_ploidy, sampGen, fuseGen, rep, driftGen)) %>% mutate(size.taj = map(data, ~getDims(., y.metric = "Tajima.D_1")), size.h = map(data, ~getDims(., y.metric = "Fay.Wu.H_1")), size.e = map(data, ~getDims(., y.metric = "Zeng.E_1")), size.fst = map(data, ~getDims(., y.metric = "Fst")), size.dxy = map(data, ~getDims(., y.metric = "dxy.1.2")), med.d = map(data, ~getStat(., "Tajima.D_1")), med.e = map(data, ~getStat(., "Zeng.E_1")), med.h = map(data, ~getStat(., "Fay.Wu.H_1")), med.fst = map(data, ~getStat(., "Fst")), med.dxy = map(data, ~getStat(., "dxy.1.2"))) %>% select(-data) %>% unnest()

#Combined sample gen figures
kkk = merge(dd, ff, by = c("traj_ploidy", "sim_ploidy", "dom", "s", "N", "mu", "recomb", "sampGen", "fuseGen", "rep"), all = T)

kkk.1 = kkk %>% filter(dom == 0.5 & traj_ploidy == sim_ploidy & s == "0.01" & N=="10000" & driftGen == 0 & recomb == 1e-08)  %>% mutate("TajD" = med.d, "FayWuH" = med.h, "ZengsE" = med.e, "Fst" = med.fst, "Dxy" = med.dxy, "XPEHH" = med.xp, "iHS" = med.ihs) %>% select(sim_ploidy, sampGen, TajD, ZengsE, FayWuH, Fst, Dxy, XPEHH, iHS) %>% melt(., id.var = c("sim_ploidy", "sampGen")) %>% mutate(stat = "MaxValue")

kkk.med.points = kkk %>% filter(dom == 0.5 & traj_ploidy == sim_ploidy & s == "0.01" & N=="10000" & driftGen == 0 & recomb == 1e-08)  %>% mutate("TajD" = med.d, "FayWuH" = med.h, "ZengsE" = med.e, "Fst" = med.fst, "Dxy" = med.dxy, "XPEHH" = med.xp, "iHS" = med.ihs) %>% select(sim_ploidy, sampGen, TajD, ZengsE, FayWuH, Fst, Dxy, XPEHH, iHS) %>% melt(., id.var = c("sim_ploidy", "sampGen")) %>% mutate(stat = "MaxValue") %>% group_by(sim_ploidy, sampGen, variable) %>% summarize(value = median(value, na.rm=T))

kkk.2 = kkk %>% filter(dom == 0.5 & traj_ploidy == sim_ploidy & s == "0.01" & N=="10000" & driftGen == 0 & recomb == 1e-08)  %>% mutate("TajD" = size.taj, "FayWuH" = size.h, "ZengsE" = size.e, "Fst" = size.fst, "Dxy" = size.dxy, "XPEHH" = size.xp, "iHS" = size.ihs) %>% select(sim_ploidy, sampGen, TajD, ZengsE, FayWuH, Fst, Dxy, XPEHH, iHS) %>% melt(., id.var = c("sim_ploidy", "sampGen")) %>% mutate(stat = "AUC")

kkk.3 = rbind(kkk.1, kkk.2)

#Combined AUC and MaxValue
kkk.3 %>% filter(variable %in% c("sim_ploidy", "sampGen", "TajD", "Fst", "iHS", "XPEHH") & stat == "AUC") %>% mutate(variable = factor(variable, levels = c("TajD", "Fst", "iHS", "XPEHH"))) %>% ggplot(., aes(y = value / 100000, fill = as.factor(sampGen), x = as.factor(sim_ploidy))) + geom_boxplot(outlier.shape = NA) + geom_point(data = kkk.med.points[kkk.med.points$variable %in% c("TajD", "Fst", "iHS", "XPEHH"),], aes(y = value), shape = 18, color = "blue", position = position_dodge(width = 0.75), show.legend = F, size=2.5)+ scale_fill_manual(name = "Generations\nsince Fixation", values = wes_palette("Royal2")) + scale_y_continuous(labels=scaleFUN) + ylab("Peak Area") + xlab("Ploidy") + facet_wrap(~variable, scales = "free", shrink = T, nrow = 1) + theme_bw() + theme(axis.text.y=element_text(size = 12), axis.title.y=element_text(size=16), axis.text.x=element_text(size=13), axis.title.x=element_text(size = 16), legend.text = element_text(size=13), legend.title = element_text(size=14), strip.text = element_text(size=14))
```

#Figure 2E
```{r}
ds = ms2 %>% group_by(sim_ploidy, DS, rep, dom) %>% mutate(Div = mean(Pi.2) - Pi.1, Tajima.D_1 = abs(Tajima.D_1), Fst = scale(fst)) %>% ungroup() %>% nest(-c(sim_ploidy, dom, DS, rep)) %>% mutate(size.div = map(data, ~getDims(., y.metric = "Div")), size.taj = map(data, ~getDims(., y.metric = "Tajima.D_1")), size.h = map(data, ~getDims(., y.metric = "Fay.Wu.H_1")), size.e = map(data, ~getDims(., y.metric = "Zeng.E_1")), size.fst = map(data, ~getDims(., y.metric = "Fst")), size.dxy = map(data, ~getDims(., y.metric = "dxy.1.2")), med.d = map(data, ~getStat(., "Tajima.D_1")), med.e = map(data, ~getStat(., "Zeng.E_1")), med.h = map(data, ~getStat(., "Fay.Wu.H_1")), med.fst = map(data, ~getStat(., "Fst")), med.dxy = map(data, ~getStat(., "dxy.1.2"))) %>% select(-data) %>% unnest()

ms2b = merge(ds, ehh, by = c("sim_ploidy", "dom", "DS", "rep"), all = T)

MS6.1 = ms2b %>% mutate("TajD" = med.d, "FayWuH" = med.h, "ZengsE" = med.e, "Fst" = med.fst, "Dxy" = med.dxy, "XPEHH" = med.xp, "iHS" = med.ihs) %>% select(sim_ploidy, DS, TajD, ZengsE, FayWuH, Fst, Dxy, XPEHH, iHS) %>% group_by(sim_ploidy, DS) %>% summarize_all(c("CV", "lCV", "uCV"))

MS6.1 %>% melt(., id.var = c("sim_ploidy", "DS")) %>% separate(variable, c("variable","XX"), "_") %>% spread(XX, value) %>% filter(variable %in% c("TajD", "Fst", "iHS", "XPEHH")) %>% mutate(Variable = factor(variable, levels=c("TajD", "Fst", "iHS", "XPEHH"))) %>% ggplot(., aes(y = CV, fill = as.factor(DS), x = as.factor(sim_ploidy))) + geom_bar(stat="identity", position = position_dodge()) + geom_errorbar(aes(ymin = lCV, ymax = uCV, width = 0.5), position = position_dodge(width = 0.9)) + scale_fill_discrete(name = "Sampling", labels=c("Equal Individuals", "Equal Alleles")) + scale_y_continuous(labels=scaleFUN) + ylab("CV[MaxValue]") + xlab("Ploidy") + facet_wrap(~Variable, scales = "free", shrink = T, nrow = 1) + theme_bw() + theme(axis.text.y=element_blank(), axis.title.y=element_text(size=16), axis.text.x=element_text(size=14), axis.title.x=element_text(size=16), legend.text = element_text(size=13), legend.title = element_text(size=14), strip.text = element_text(size=14))
```

Supp Fig 5
```{r}

rehh2 = read.csv(paste("~/Documents/Research/PloidySim/rehh/", "rehh_calcs_do0.1_do0.9_sG_nDS.csv", sep=""))

rehh2 %<>% mutate(driftGen = 0) %>% select(-X)

rehh3 = read.csv(paste("~/Documents/Research/PloidySim/rehh/", "rehh_calcs_do0.1_sG_mkup3.csv", sep=""))

rehh3 %<>% mutate(driftGen = 0)

rehh2 = rbind(rehh2,rehh3)

ff2 = rehh2 %>% filter(!is.na(xpehh.p) & sampGen!=1000) %>% mutate(mid = POSITION) %>% nest(-c(s, dom, recomb, N, mu, sim_ploidy, traj_ploidy, sampGen, fuseGen, driftGen, rep)) %>% mutate(size.xp = map(data, ~getDims(., y.metric = "XPEHH")), size.xp.p = map(data, ~getDims(., y.metric = "xpehh.p")), med.xp = map(data, ~getStat(., "XPEHH")), med.xpp = map(data, ~log(getStat(., "xpehh.p"))), size.ihs = map(data, ~getDims(., y.metric = "IHS")), size.ihs.p = map(data, ~getDims(., y.metric = "ihs.p")), med.ihs = map(data, ~getStat(., "IHS")), med.ihs.p = map(data, ~log(getStat(., "ihs.p")))) %>% select(-c(data)) %>% unnest()

ff2 %<>% mutate(traj_ploidy = as.numeric(str_replace(traj_ploidy, "tp","")), sim_ploidy = as.numeric(str_replace(sim_ploidy, "sp","")), dom = as.numeric(str_replace(dom, "do","")), s = as.numeric(str_replace(s, "se","")), N = as.numeric(str_replace(N, "NN","")), mu = as.numeric(str_replace(mu, "mu","")), recomb = as.numeric(str_replace(recomb, "re","")), sampGen = as.numeric(str_replace(sampGen, "sG","")), fuseGen = as.numeric(str_replace(fuseGen, "fG","")), rep = as.numeric(str_replace(rep, "rep", "")))

kkk2 = merge(dd, ff2, by = c("traj_ploidy", "sim_ploidy", "dom", "s", "N", "mu", "recomb", "sampGen", "fuseGen", "rep", "driftGen"), all = T)

#Recessive
kkk.1b = kkk2 %>% filter(dom == 0.1 & traj_ploidy == sim_ploidy & s == "0.01" & N=="10000" & driftGen == 0 & recomb == 1e-08)  %>% mutate("TajD" = med.d, "FayWuH" = med.h, "ZengsE" = med.e, "Fst" = med.fst, "Dxy" = med.dxy, "XPEHH" = med.xp, "iHS" = med.ihs) %>% select(sim_ploidy, sampGen, TajD, ZengsE, FayWuH, Fst, Dxy, XPEHH, iHS) %>% melt(., id.var = c("sim_ploidy", "sampGen")) %>% mutate(stat = "MaxValue")

kkk.med.points.b = kkk2 %>% filter(dom == 0.1 & traj_ploidy == sim_ploidy & s == "0.01" & N=="10000" & driftGen == 0 & recomb == 1e-08)  %>% mutate("TajD" = med.d, "FayWuH" = med.h, "ZengsE" = med.e, "Fst" = med.fst, "Dxy" = med.dxy, "XPEHH" = med.xp, "iHS" = med.ihs) %>% select(sim_ploidy, sampGen, TajD, ZengsE, FayWuH, Fst, Dxy, XPEHH, iHS) %>% melt(., id.var = c("sim_ploidy", "sampGen")) %>% mutate(stat = "MaxValue") %>% group_by(sim_ploidy, sampGen, variable) %>% summarize(value = median(value, na.rm=T))

kkk.2b = kkk2 %>% filter(dom == 0.1 & traj_ploidy == sim_ploidy & s == "0.01" & N=="10000" & driftGen == 0 & recomb == 1e-08)  %>% mutate("TajD" = size.taj, "FayWuH" = size.h, "ZengsE" = size.e, "Fst" = size.fst, "Dxy" = size.dxy, "XPEHH" = size.xp, "iHS" = size.ihs) %>% select(sim_ploidy, sampGen, TajD, ZengsE, FayWuH, Fst, Dxy, XPEHH, iHS) %>% melt(., id.var = c("sim_ploidy", "sampGen")) %>% mutate(stat = "AUC")

kkk.3b = rbind(kkk.1b, kkk.2b)

#Combined AUC and MaxValue
kkk.3b %>% filter(variable %in% c("sim_ploidy", "sampGen", "TajD", "Fst", "iHS", "XPEHH") & stat == "AUC") %>% mutate(variable = factor(variable, levels = c("TajD", "Fst", "iHS", "XPEHH"))) %>% ggplot(., aes(y = value / 100000, fill = as.factor(sampGen), x = as.factor(sim_ploidy))) + geom_boxplot(outlier.shape = NA) + geom_point(data = kkk.med.points.b[kkk.med.points.b$variable %in% c("TajD", "Fst", "iHS", "XPEHH"),], aes(y = value), shape = 18, color = "blue", position = position_dodge(width = 0.75), show.legend = F, size=2.5)+ scale_fill_manual(name = "Generations\nsince Fixation", values = wes_palette("Royal2")) + scale_y_continuous(labels=scaleFUN) + ylab("Peak Area") + xlab("Ploidy") + facet_wrap(~variable, scales = "free", shrink = T, nrow = 1) + theme_bw() + theme(axis.text.y=element_text(size = 12), axis.title.y=element_text(size=16), axis.text.x=element_text(size=13), axis.title.x=element_text(size = 16), legend.text = element_text(size=13), legend.title = element_text(size=14), strip.text = element_text(size=14))

#Dominant
kkk.1c = kkk2 %>% filter(dom == 0.9 & traj_ploidy == sim_ploidy & s == "0.01" & N=="10000" & driftGen == 0 & recomb == 1e-08)  %>% mutate("TajD" = med.d, "FayWuH" = med.h, "ZengsE" = med.e, "Fst" = med.fst, "Dxy" = med.dxy, "XPEHH" = med.xp, "iHS" = med.ihs) %>% select(sim_ploidy, sampGen, TajD, ZengsE, FayWuH, Fst, Dxy, XPEHH, iHS) %>% melt(., id.var = c("sim_ploidy", "sampGen")) %>% mutate(stat = "MaxValue")

kkk.med.points.c = kkk2 %>% filter(dom == 0.9 & traj_ploidy == sim_ploidy & s == "0.01" & N=="10000" & driftGen == 0 & recomb == 1e-08)  %>% mutate("TajD" = med.d, "FayWuH" = med.h, "ZengsE" = med.e, "Fst" = med.fst, "Dxy" = med.dxy, "XPEHH" = med.xp, "iHS" = med.ihs) %>% select(sim_ploidy, sampGen, TajD, ZengsE, FayWuH, Fst, Dxy, XPEHH, iHS) %>% melt(., id.var = c("sim_ploidy", "sampGen")) %>% mutate(stat = "MaxValue") %>% group_by(sim_ploidy, sampGen, variable) %>% summarize(value = median(value, na.rm=T))

kkk.2c = kkk2 %>% filter(dom == 0.9 & traj_ploidy == sim_ploidy & s == "0.01" & N=="10000" & driftGen == 0 & recomb == 1e-08)  %>% mutate("TajD" = size.taj, "FayWuH" = size.h, "ZengsE" = size.e, "Fst" = size.fst, "Dxy" = size.dxy, "XPEHH" = size.xp, "iHS" = size.ihs) %>% select(sim_ploidy, sampGen, TajD, ZengsE, FayWuH, Fst, Dxy, XPEHH, iHS) %>% melt(., id.var = c("sim_ploidy", "sampGen")) %>% mutate(stat = "AUC")

kkk.3c = rbind(kkk.1c, kkk.2c)

#Combined AUC and MaxValue
kkk.3c %>% filter(variable %in% c("sim_ploidy", "sampGen", "TajD", "Fst", "iHS", "XPEHH") & stat == "AUC") %>% mutate(variable = factor(variable, levels = c("TajD", "Fst", "iHS", "XPEHH"))) %>% ggplot(., aes(y = value / 100000, fill = as.factor(sampGen), x = as.factor(sim_ploidy))) + geom_boxplot(outlier.shape = NA) + geom_point(data = kkk.med.points.c[kkk.med.points.c$variable %in% c("TajD", "Fst", "iHS", "XPEHH"),], aes(y = value), shape = 18, color = "blue", position = position_dodge(width = 0.75), show.legend = F, size=2.5)+ scale_fill_manual(name = "Generations\nsince Fixation", values = wes_palette("Royal2")) + scale_y_continuous(labels=scaleFUN) + ylab("Peak Area") + xlab("Ploidy") + facet_wrap(~variable, scales = "free", shrink = T, nrow = 1) + theme_bw() + theme(axis.text.y=element_text(size = 12), axis.title.y=element_text(size=16), axis.text.x=element_text(size=13), axis.title.x=element_text(size = 16), legend.text = element_text(size=13), legend.title = element_text(size=14), strip.text = element_text(size=14))
```

Supplemental Figure 
```{r}

lCV <- function(data, N = 1000, n = 50){
    aa = foreach(i = 1:N, .combine = rbind) %do% {
      ndat = sample(data, n)
      sd(ndat, na.rm=T)/(mean(ndat, na.rm=T))*100}
    return(quantile(aa,0.05, na.rm=T))
}

uCV <- function(data, N = 1000, n = 50){
    aa = foreach(i = 1:N, .combine = rbind) %do% {
      ndat = sample(data, n)
      sd(ndat, na.rm=T)/(mean(ndat, na.rm=T))*100}
    return(quantile(aa,0.95, na.rm=T))
}

pardir = "..." #Set directory housing data for SuppFig6
ehh = read.csv(paste(pardir, "rehh_calcs_SuppFig6.csv", sep = ""))

ds2 = ms2 %>% group_by(sim_ploidy, DS, rep, dom) %>% mutate(Div = mean(Pi.2) - Pi.1, Tajima.D_1 = abs(Tajima.D_1), Fst = scale(fst)) %>% ungroup() %>% nest(-c(sim_ploidy, dom, DS, rep)) %>% mutate(size.div = map(data, ~getDims(., y.metric = "Div")), size.taj = map(data, ~getDims(., y.metric = "Tajima.D_1")), size.h = map(data, ~getDims(., y.metric = "Fay.Wu.H_1")), size.e = map(data, ~getDims(., y.metric = "Zeng.E_1")), size.fst = map(data, ~getDims(., y.metric = "Fst")), size.dxy = map(data, ~getDims(., y.metric = "dxy.1.2")), med.d = map(data, ~getStat(., "Tajima.D_1")), med.e = map(data, ~getStat(., "Zeng.E_1")), med.h = map(data, ~getStat(., "Fay.Wu.H_1")), med.fst = map(data, ~getStat(., "Fst")), med.dxy = map(data, ~getStat(., "dxy.1.2"))) %>% select(-data) %>% unnest()

ms3 = merge(ds2, ehh2, by = c("sim_ploidy", "DS", "rep", "dom"), all = T)

MS6.1b = ms3 %>% filter(dom==0.9) %>%  mutate("TajD" = med.d, "FayWuH" = med.h, "ZengsE" = med.e, "Fst" = med.fst, "Dxy" = med.dxy, "XPEHH" = med.xp, "iHS" = med.ihs) %>% select(sim_ploidy, dom, DS, TajD, ZengsE, FayWuH, Fst, Dxy, XPEHH, iHS) %>% group_by(sim_ploidy, dom, DS) %>% summarize_all(c("CV", "lCV", "uCV"))

MS6.1b = ms3 %>% mutate("TajD" = med.d, "FayWuH" = med.h, "ZengsE" = med.e, "Fst" = med.fst, "Dxy" = med.dxy, "XPEHH" = med.xp, "iHS" = med.ihs) %>% select(sim_ploidy, dom, DS, TajD, ZengsE, FayWuH, Fst, Dxy, XPEHH, iHS) %>% group_by(sim_ploidy, dom, DS) %>% summarize_all(c("CV", "lCV", "uCV"))

MS6.1b %>% filter(dom==0.1) %>% melt(., id.var = c("sim_ploidy", "DS")) %>% separate(variable, c("variable","XX"), "_") %>% spread(XX, value) %>% filter(variable %in% c("TajD", "Fst", "iHS", "XPEHH")) %>% mutate(Variable = factor(variable, levels=c("TajD", "Fst", "iHS", "XPEHH"))) %>% ggplot(., aes(y = CV, fill = as.factor(DS), x = as.factor(sim_ploidy))) + geom_bar(stat="identity", position = position_dodge()) + geom_errorbar(aes(ymin = lCV, ymax = uCV, width = 0.5), position = position_dodge(width = 0.9)) + scale_fill_discrete(name = "Sampling", labels=c("Equal Individuals", "Equal Alleles")) + scale_y_continuous(labels=scaleFUN) + ylab("CV[MaxValue]") + xlab("Ploidy") + facet_wrap(~Variable, scales = "free", shrink = T, nrow = 1) + theme_bw() + theme(axis.text.y=element_blank(), axis.title.y=element_text(size=16), axis.text.x=element_text(size=14), axis.title.x=element_text(size=16), legend.text = element_text(size=13), legend.title = element_text(size=14), strip.text = element_text(size=14))
MS6.1b %>% filter(dom==0.9) %>% melt(., id.var = c("sim_ploidy", "DS")) %>% separate(variable, c("variable","XX"), "_") %>% spread(XX, value) %>% filter(variable %in% c("TajD", "Fst", "iHS", "XPEHH")) %>% mutate(Variable = factor(variable, levels=c("TajD", "Fst", "iHS", "XPEHH"))) %>% ggplot(., aes(y = CV, fill = as.factor(DS), x = as.factor(sim_ploidy))) + geom_bar(stat="identity", position = position_dodge()) + geom_errorbar(aes(ymin = lCV, ymax = uCV, width = 0.5), position = position_dodge(width = 0.9)) + scale_fill_discrete(name = "Sampling", labels=c("Equal Individuals", "Equal Alleles")) + scale_y_continuous(labels=scaleFUN) + ylab("CV[MaxValue]") + xlab("Ploidy") + facet_wrap(~Variable, scales = "free", shrink = T, nrow = 1) + theme_bw() + theme(axis.text.y=element_blank(), axis.title.y=element_text(size=16), axis.text.x=element_text(size=14), axis.title.x=element_text(size=16), legend.text = element_text(size=13), legend.title = element_text(size=14), strip.text = element_text(size=14))
```